---
title: "vizWithSCE"
author: "Kwame Forbes"
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vizWithSCE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

*vizWithSCE* is an R package that assists with the visualization of
the integration of bulk DE results tables with pre-processed scRNA-seq
datasets available on Bioconductor. vizWithSCE currently takes two
arguments, a list containing res, dds, and a SingleCellExperiment
(sce) as selected by the user and a second argument which. The which
argument is the gene to use for the violin plot, represented as a
number that corresponds to the gene with the n-th smallest 
p-value, e.g., 1 for the gene with lowest p-value, etc. 

# Run DESeq2 to obtain bulk DE results

We first load gene-summarized RNA-seq counts.

```{r eval=FALSE}
library(airway)
```

For our workflow, we used the bulk dataset "airway", which contains
four human airway smooth muscle cell lines treated with dexamethasone,
a steroid used to treat inflammation. This pre-processed dataset is
available on Bioconductor. 
Here we filter to at least 3 samples with a count of 10 or higher, and
then perform differential expression analysis.

Mike: why did you change the rownames?
To match with the format?

```{r eval=FALSE}
library(DESeq2)
data(gse)
levels(gse$condition) <- c("untrt", "trt")
dds <- DESeqDataSet(gse, design = ~donor + condition)
keep <- rowSums(counts(dds) >= 10) >= 3
dds <- dds[keep,]
dds <- DESeq(dds)
res <- results(dds)
rownames(res) <- sub("\\..*","",rownames(res))
```

```{r echo=FALSE, results="hide"}
# the following is evaluated but hidden,
# because the above un-evaluated code chunk
# is slow to run
library(vizWithSCE)
data(dds)
dds <- DESeq2::estimateSizeFactors(dds)
data(res)
```

# Run `integrateWithSingleCell`

We run `integrateWithSingleCell` and pick number 8 from the dataset
selection as our single-cell dataset. Within PBMC there are nine
different scRNA-seq datasets. We chose the dataset pbmc4k, which
contains roughly 4 thousand cells. This pre-processed dataset is
available on Bioconductor. 

```{r eval=FALSE}
dat <- integrateWithSingleCell(res, dds)
```

```{r echo=FALSE, results="hide", message=FALSE}
# the following is evaluated but hidden,
# because the above un-evaluated code chunk
# involves user input.
library(TENxPBMCData)
library(scater)
tenx_pbmc4k <- TENxPBMCData(dataset = "pbmc4k")
sce <- tenx_pbmc4k
dat <- list(res=res, dds=dds, sce=sce)
dat$sce <- logNormCounts(dat$sce)
```

# Label cell types or clusters

Mike: what is scater used here for?
Used to add logNormCounts to sce?

```{r eval=FALSE}
library(scater)
dat$sce <- scater::logNormCounts(dat$sce)
sce2 <- dat$sce
```

Here we convert the rownames from Ensembl to symbol, to match with the
cell annotation reference dataset, described below.

```{r eval=FALSE}
library(org.Hs.eg.db)
rownames(sce2) <- mapIds(org.Hs.eg.db, rownames(sce2), "SYMBOL", "ENSEMBL")
```

Mike: what is celldex used for here? what is SingleR used here for?
celldex database of reference expression datasets 
SingleR cell type recognition
Mike: after this step what has changed in `dat`?
dat now contains labels of different cell types

```{r eval=FALSE}
library(celldex)
library(SingleR)
ref <- celldex::BlueprintEncodeData()
pred <- SingleR::SingleR(test=sce2, ref=ref, labels=ref$label.main)
table(pred$labels)
colLabels(dat$sce) <- pred$labels
```

```{r echo=FALSE, results="hide"}
# the following is evaluated but hidden,
# because the above un-evaluated code chunk
# is slow to run
data(labels)
colLabels(dat$sce) <- labels
```

# Use *vizWithSCE* for visualization 

Mike: what does the following plot tell us about the gene with the 2nd
smallest p-value? Which single cells is it expressed in?
DUSP1,known to play a role in response to environmental stress, is most 
highly expressed in monocytes

```{r viz-plot}
library(vizWithSCE)
vizWithSCE(dat, which=2)
```

# Future direction

Mike: what else would you like to do with the package? You can list
anything you like here.
We are currently planning to add more features, including other types of plots,
allowing users to further explore the intersection of the two datasets.
We seek to add multiple gene plots, e.g. dotplots or heatmaps of sets of DE 
genes.
We are considering statistical tests of the distribution of expression across 
cell types
for sets of DE genes, e.g. up-regulated vs down-regulated DE genes.
We will explore launching iSEE from our worklow, a well-developed package for
interactive exploration of SCE datasets.

# Session info

```{r}
sessionInfo()
```
